---
title: "EDDA_asg3_3"
author: "Chenghan Song"
date: "2020/3/15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise 3

##### a)

```{r}
afrdata=read.table(file="africa.txt",header=TRUE);
pairs(afrdata);
afrdataglm=glm(miltcoup~oligarchy+pollib+parties+pctvote+popn+size+numelec+numregim, family=poisson,data=afrdata);
summary(afrdataglm);
```
```{r}
confint(afrdataglm)
```


$H_0$: The explanatory variable does not have a main effect on the number of successful military coups from independence to 1989.

The p-value for oligarchy is 0.035 meaning that we can safely reject $H_0$ concluding that number years country ruled by military oligarchy from independence to 1989 has a significant effect on the number of successful military coups from independence to 1989. We are 95% confident that the mean number of successful military coups from independence to 1989 increases between 0.5% and 14% for each additional year country ruled by military oligarchy.

The p-value for pollib is 0.0089 meaning that we reject $H_0$ concluding that political liberalization has an effect on the number of successful military coups from independence to 1989. We are 95% confident that the mean number of successful military coups from independence to 1989 decreases between 18% and 126% for each level the political liberalization increases. 

The p-value for parties is 0.00595 meaning that we can safely reject $H_0$ concluding that number of legal political parties in 1993 has a significant effect on the number of successful military coups from independence to 1989. We are 95% confident that the mean number of successful military coups from independence to 1989 increases between 0.8% and 5% for each additional increase of legal political parties in 1993.

```{r}
par(mfrow=c(2,2));
plot(fitted(afrdataglm),residuals(afrdataglm),main="residuals VS fitted values");plot(log(fitted(afrdataglm)),residuals(afrdataglm),main="residuals VS log fitted values");
plot(log(fitted(afrdataglm)),residuals(afrdataglm,type="response"),main="response residuals VS fitted values")
```

From the plots above we can see some specific patterns which means the model does not fit in this case. 


##### b)
```{r}
summary(glm(miltcoup~oligarchy+pollib+parties+pctvote+popn+size+numelec+numregim,data=afrdata))[12]
```

```{r}
summary(glm(miltcoup~oligarchy+pollib+parties+pctvote+popn+size+numelec,data=afrdata))[12]
```
```{r}
summary(glm(miltcoup~oligarchy+pollib+parties+pctvote+popn+size,data=afrdata))[12]
```
```{r}
summary(glm(miltcoup~oligarchy+pollib+parties+popn+size,data=afrdata))[12]
```
```{r}
summary(glm(miltcoup~oligarchy+pollib+parties+popn,data=afrdata))[12]
```

```{r}
summary(glm(miltcoup~oligarchy+pollib+parties,data=afrdata))
```
By using the step down approach, we find that oligarchy, pollib, and parties are the essential in explaining the number of successful military coups from independence to 1989. Other explanatory variables:pctvote, popn, size, numelec, and numregim should be removed.

The resulting model of the step-down method is: miltcoup = 1.62597 + 0.15261 $*$ oligarchy - 0.95419 $*$ pollib + 0.04239 $*$ parties + error


