---
title: "edda_final"
author: "Chenghan Song"
date: "2020/3/24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise Tree

##### a)

```{r}
forest=read.table(file="treeVolume.txt",header=TRUE);
```

##### b)

##### c)

## Exercise Jane Austen

##### a)
The interest of contingency table test is in a possible dependency of the two factors. Thus it is suitable and appropriate for studying possible dependency between Austen's novels and an Austen's admirer's novel(two factors) by counting the number of different words(categories). 

##### b)

```{r}
book =read.table(file="austen.txt",header=TRUE);
ausmatrix = as.matrix(book[,1:3]);
auscq=chisq.test(ausmatrix);
auscq
(auscq$observed-auscq$expected)/sqrt(auscq$expected);
```
$H_0$: There is no difference between the words counts/literary styles in Austen's three books.
The p-value is 0.2673 meaning that we do not reject $H_0$ concluding that there is no difference between the words counts/literary styles with different books. Austen herself was consistent in her different novels.

##### c)

```{r}
book =read.table(file="austen.txt",header=TRUE);
bookmatrix = as.matrix(book);
z=chisq.test(bookmatrix);
z
(z$observed-z$expected)/sqrt(z$expected);
```
$H_0$: There is no difference between the words counts/literary styles in different books.

The p-value is 6.205e-05 meaning that we can reject $H_0$ concluding that there is significant difference between the words counts/literary styles with different books. The admiror was not successful in imitating Austen's style. From the table we can see the admirer used relatively more 'an','with' and less 'that','without' in Sanditon while Austen used relatively more 'a','without' and less 'an','with' in Sanditon, which made the whole novel inconsist. 

